---
title: European Energy Data Analysis
author: ~
date: '2020-12-08'
slug: european-energy-data-analysis
categories: []
tags: []
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(ggplot2)
library(tidyverse)
library(readxl)
library(ggthemes)
library(patchwork)

energy_types <- readr::read_csv(here::here("data", "tidytuesday", "energy_types.csv"))
country_totals <- readr::read_csv(here::here("data", "tidytuesday", "country_totals.csv"))
```

# European Energy Usage: Clean Energy vs. Conventional

Using data that is available on the Tidy Tuesday website on European Energy, I will investigate the energy usage among European countries with an emphasis on which countries are the best at using renewable energy. I will be examining data from the `energy_types.csv` file from the Tidy Tuesday website. The data consists of 296 observations and 7 variables. The data displays each the energy production from each European country from 2016-2018 and breaks it down by how much energy was produced by conventional thermal, nuclear, hydroelectric, pumped hydroelectric, wind, solar, geothermal, and other.

The second dataset I will use to explore European Energy production is the `country_totals.csv` file from the Tidy Tuesday website. The data consists of 185 observations and 7 variables. The data displays the total energy consumption of each European country from 
2016-2018 as well as the total energy production of each country.

# Question 1

Which European country has the largest percentage of their total energy production made up of renewables? To find this, I joined the energy types and country totals datasets. I will also mutate the dataset to combine all of types of energy production that are renewable.

```{r}
energy_types %>%
  left_join(country_totals, by = c("country", "country_name"))
```