---
title: European Energy Data Analysis
author: ~
date: '2020-12-08'
slug: european-energy-data-analysis
categories: []
tags: []
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(ggplot2)
library(tidyverse)
library(readxl)
library(ggthemes)
library(patchwork)

energy_types <- readr::read_csv(here::here("data", "tidytuesday", "energy_types.csv"))
country_totals <- readr::read_csv(here::here("data", "tidytuesday", "country_totals.csv"))
```

# European Energy Usage: Clean Energy vs. Conventional

Using data that is available on the Tidy Tuesday website on European Energy, I will investigate the energy usage among European countries with an emphasis on which countries are the best at using renewable energy and clean energy. I will be examining data from the `energy_types.csv` file from the Tidy Tuesday website. The data consists of 296 observations and 7 variables. The data displays the energy production from each European country from 2016-2018 and breaks it down by how much energy was produced by conventional thermal, nuclear, hydroelectric, pumped hydroelectric, wind, solar, geothermal, and other.

The second dataset I will use to explore European Energy production is the `country_totals.csv` file from the Tidy Tuesday website. The data consists of 185 observations and 7 variables. The data displays the total energy consumption of each European country from 2016-2018 as well as the total energy production of each country.

For the purposes of this blog post Renewable Energy will be hydroelectric, pumped hydroelectric, wind, solar, geothermal, and other energy productions and clean energy will be renewable energy plus nuclear energy.

# Question 1

Which European country had the largest percentage of their total energy production made up of renewable energy in 2018? Which European country had the largest percentage of their total energy production made up of clean energy? To find this, I made 3 separate data frames to summarize the total renewable energy, nuclear energy, and conventional thermal energy that each European Country produced. I then joined the 3 data frames and selected the columns with the 2018 data to make one dataset with all of the summarized energy data from 2018.

```{r, message=FALSE}
dat <- energy_types %>%
  left_join(country_totals, by = c("country", "country_name")) %>%
  rename(twenty_16.x = '2016.x', twenty_17.x = '2017.x', twenty_18.x = '2018.x') %>%
  filter(type.x %in% c("Hydro", "Pumped hydro power", "Wind", "Solar", "Geothermal", "Other")) %>%
  group_by(country) %>%
  summarize(renew_2016 = sum(twenty_16.x), renew_2017 = sum(twenty_17.x), renew_2018 = sum(twenty_18.x)) 

dat2 <-  energy_types %>%
  filter(type == "Conventional thermal") %>%
  rename(twenty_16 = '2016', twenty_17 = '2017', twenty_18 = '2018') %>%
  group_by(country) %>%
  summarize(conventional_2016 = sum(twenty_16), conventional_2017 = sum(twenty_17), conventional_2018 = sum(twenty_18))
  
dat3 <-  energy_types %>%
  filter(type == "Nuclear") %>%
  rename(twenty_16 = '2016', twenty_17 = '2017', twenty_18 = '2018') %>%
  group_by(country) %>%
  summarize(nuclear_2016 = sum(twenty_16), nuclear_2017 = sum(twenty_17), nuclear_2018 = sum(twenty_18))

data_18 <- country_totals  %>%
  left_join(dat, by = "country") %>%
  left_join(dat2, by = "country") %>%
  left_join(dat3, by = "country") %>%
  filter(type == "Total net production") %>%
  select(1, 2, 10, 13 , 16) %>%
  mutate(total_2018 = renew_2018 + conventional_2018 + nuclear_2018)
```

Next I used the dataset that I made with the summarized 2018 energy data to make 2 bar graphs that compare all the European Countries based on the percentage of their total energy output was made up of Renewable Energy and of Clean Energy, respectively, in 2018.

```{r}
ggplot(data_18, aes(x = reorder(country, -renew_2018/total_2018), y = renew_2018/total_2018)) +
  geom_bar(stat = "identity") +
  scale_y_continuous(labels = scales::percent) +
  labs(x = "Country", y = "Percentage Renewable Energy") +
  ggtitle("Percentage of Total Energy Production that was Renewable 
          by European Country in 2018") +
  theme_economist() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  theme(plot.title = element_text(size = 14, hjust = 0.35))

ggplot(data_18, aes(x = reorder(country, -(renew_2018 + nuclear_2018)/total_2018), y = (renew_2018 + nuclear_2018)/total_2018)) +
  geom_bar(stat = "identity") +
  scale_y_continuous(labels = scales::percent) +
  labs(x = "Country", y = "Percentage Clean Energy") +
  ggtitle("Percentage of Total Energy Production that was Clean 
          by European Country in 2018") +
  theme_economist() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  theme(plot.title = element_text(size = 14, hjust = 0.35))
```
Albania was the country with both the highest percentage of Renewable energy and Clean energy, since 100% of their energy production was made by renewable sources. Norway was not far behind with over 90% of their energy production coming from renewable sources. Also Countries such as France and Sweden saw a large increase when comparing their Renewable Energy percentage to their Clean Energy percentage, indicating a large use of nuclear power in these countries.

# Question 2

Which European country made the biggest shift towards renewable energy from 2016 to 2018? To find this, I joined the 3 data frames that I already made to summarize the total renewable energy, nuclear energy, and conventional thermal energy that each European Country produced. Next, I used summarize to find the percent of energy production that was renewable in each country in both 2016 and 2018. Lastly, I used mutate to make a shift variable that shows the change in percentage of energy production that was renewable in each country from 2016 to 2018 and I used arrange to have the countries with largest shifts at the top of the data.

```{r}
shift_data <- dat  %>%
  left_join(dat2, by = "country") %>%
  left_join(dat3, by = "country") %>%
  mutate(total_2018 = renew_2018 + conventional_2018 + nuclear_2018, total_2017 = renew_2017 + conventional_2017 + nuclear_2017, total_2016 = renew_2016 + conventional_2016 + nuclear_2016) %>%
  select(1:4, 11:13) %>%
  group_by(country) %>%
  summarize(renew_16_prc = renew_2016/total_2016, renew_18_prc = renew_2018/total_2018) %>%
  mutate(shift = renew_18_prc - renew_16_prc) %>%
  arrange(desc(shift))

big_shift <- shift_data[1:10, ]
big_shift
```

Next, I shortened the data to have just the 10 European countries with the largest shifts. I then used this data to make a bar graph with these 10 counties and their percentage shift towards renewable energy, so that the results could be better visualized.

```{r}
ggplot(big_shift, aes(x = reorder(country, -shift), y = shift)) +
  geom_col() +
  scale_y_continuous(labels = scales::percent) +
  labs(x = "Country", y = "Percentage Shift to Renewable Energy") +
  ggtitle("The 10 European Countries with the biggest shift towards 
          Renewable Energy from 2016 to 2018") +
  theme_economist() +
  theme(plot.title = element_text(size = 14, hjust = 0.35))
```
Bosnia & Herzegovina was the country that saw the biggest shift, but they were followed closely by Lithuania. They were the only 2 countries that had over a 10% shift in production of renewable sources of energy.

# Question 3

Is there a correlation between the European countries with the biggest energy outputs and using more conventional thermal energy? To find this, I again joined the 3 data frames that I already made to summarize the total renewable energy, nuclear energy, and conventional thermal energy that each European Country produced. Next, I arranged the data so that the countries with the highest total energy production were at the top of the data and I made a second dataset that shortened the data down to just the 10 European countries with the highest total energy production in 2018.

```{r}
sort_total <- country_totals  %>%
  left_join(dat, by = "country") %>%
  left_join(dat2, by = "country") %>%
  left_join(dat3, by = "country") %>%
  filter(type == "Total net production") %>%
  select(1, 2, 10, 13 , 16) %>%
  mutate(total_2018 = renew_2018 + conventional_2018 + nuclear_2018, average_prc_conventional = mean(conventional_2018)/mean(total_2018)) %>%
  arrange(desc(total_2018))

top_output <- sort_total[1:10, ]
```

I used this data to make some visuals that help reveal whether or not their is a trend between the countries with the highest total energy production and a higher use of conventional thermal energy. The first graph plots all the European countries as a point based on their total energy production and their conventional energy production. The second graph has just the top 10 energy producing European nations and plots the percentage of their total energy production that is from conventional thermal sources, and I added a red dashed line that represents the average percent among all the European nations in 2018.

```{r}
ggplot(sort_total, aes(x = total_2018, y = conventional_2018)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(x = "Total Energy Production", y = "Conventional Thermal Energy Production") +
  ggtitle("Total Energy Production vs Conventional Thermal Energy
          Production in Europe for 2018") +
  theme_economist() +
  theme(plot.title = element_text(size = 14, hjust = 0.35))

ggplot(top_output, aes(x = reorder(country, -total_2018), y = conventional_2018/total_2018)) +
  geom_col() +
  scale_y_continuous(labels = scales::percent) +
  geom_hline(yintercept = 0.4499, linetype="dashed", color = "red") +
  labs(x = "Country", y = "Percentage Conventional Thermal Energy Production") +
  ggtitle("Percentage Conventional Thermal Energy Production among the 
          top 10 Energy Producing European Countries for 2018") +
  theme_economist() +
  theme(plot.title = element_text(size = 14, hjust = 0.35))
```

The data suggests that their is some trend between countries with higher total energy output and higher use of conventional thermal sources. However, there are some very notable exceptions, such as France, Sweden, and Norway.

# Conclusions:

In conclusion, the average European country gets about half of their energy production from clean sources. However, the range stretches from Albania that gets 100% of their energy from clean sources to Malta that gets 0% of their energy from clean sources. Furthermore, overall there has been a trend among European Countries towards Renewable sources of energy from 2016-2018, with some countries, such as Bosnia & Herzegovina and Lithuania, seeing over a 10% shift towards Renewable energy sources in that span. Lastly, among the larger energy producing countries in Europe there was a trend showing higher Conventional Thermal Energy usage in these countries, with a few notable exceptions, such as France, Sweden, and Norway.